---
title: "Analysis of the Difference in Means of Systolic Blood Pressure between Smokers and Non-Smokers Using  Bootstrapping"
author: "Brock Akerman and Hanan Ali"
date: "ST502 - Spring 2022"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(readr)
library(qqplotr)
library(janitor)
```
```{r, echo=FALSE, warning=FALSE, include=FALSE}
framingham_data <- read_csv("framingham_data.csv")
View(framingham_data)
attach(framingham_data)
names(framingham_data)
class(currentSmoker)
class(sysBP)
```
```{r, echo=FALSE, warning=FALSE, include=FALSE}
Q1 <- quantile(framingham_data$sysBP, .25)
Q3 <- quantile(framingham_data$sysBP, .75)
IQR <- IQR(framingham_data$sysBP)
framingham_data_no_outlier <- subset(framingham_data, framingham_data$sysBP > (Q1 -1.5*IQR) & framingham_data$sysBP < (Q3 + 1.5*IQR))
```

```{r}

```




An effective means of estimating a population parameter is the bootstrapping method.  To model the Framingham Heart Study data and their means, we will draw samples from the study and make inference based on those samples.  Computer software does the heavy lifting for this exercise as the iterations require randomization of normality and many calculations. 

The first thought when considering the use of iterative functions was the creation of a "for" loop.  We wanted to process our data efficiently and creating a loop helped cut down on much of the manual formula making.  There are predefined parameter assumptions to consider which added a nesting feature within our "for" loop.  

We considered variances of $\sigma^2_{1}$ = 1,4,9 and $\sigma^2_{2}$ = 1, samples sizes of $n_{1}$ = 10, 30, 70 and $n_{2}$ = 10, 30, 70 There are 135 combinations of 

Using R software, we randomly generated 100 data sets 
 
Codeblock.  
```{r}
N1 <- c(10,30,70) #Sampled number of samples from non-smokers, Column_1
N2 <- c(10,30,70) #Sampled number of samples from smokers, Column_2
S1 <- c(1, 4, 9) #Variance, Column_3
MU2 <- c(5, 10, 6, 4, 0) #MU1 is always same cause Mu1 is 5 so I set different values of Mu2 to get this to work: Ho:u1-u2=0 and Ha:u1-u2 is not equal to 0, Column_4, vector c(1)= equal means, c(2:5) = unequal means. 

B = 100  #Replicating 100 times
alpha = 0.05 #Significance level.
result1 = matrix(0, nr=135, nc=B)  #Creation of a dummy matrix to record p-values for equal variances
result2 = matrix(0, nr=135, nc=B)  #Creation of a dummy matrix to record p-values for unequal variances

para = matrix(0, nr=135, nc=4)  #Creation of a dummy matrix to record our four parameters (N1, N2, S1, MU2)
#The following functions on "para" are meant to populate the matrice.  
colnames(para) <- c("N1", "N2", "S1", "MU2") #Names the columns in the para matrix. 
para[1:5,1]=10;     para[1:5,2]=10;     para[1:5,3]=1;      para[1:5,4]=MU2   
para[6:10,1]=10;    para[6:10,2]=30;    para[6:10,3]=1;     para[6:10,4]=MU2
para[11:15,1]=10;   para[11:15,2]=70;   para[11:15,3]=1;    para[11:15,4]=MU2
para[16:20,1]=10;   para[16:20,2]=10;   para[16:20,3]=4;    para[16:20,4]=MU2
para[21:25,1]=10;   para[21:25,2]=30;   para[21:25,3]=4;    para[21:25,4]=MU2
para[26:30,1]=10;   para[26:30,2]=70;   para[26:30,3]=4;    para[26:30,4]=MU2
para[31:35,1]=10;   para[31:35,2]=10;   para[31:35,3]=9;   para[31:35,4]=MU2
para[36:40,1]=10;   para[36:40,2]=30;   para[36:40,3]=9;   para[36:40,4]=MU2
para[41:45,1]=10;   para[41:45,2]=70;   para[41:45,3]=9;   para[41:45,4]=MU2
para[46:50,1]=30;   para[46:50,2]=10;   para[46:50,3]=1;    para[46:50,4]=MU2  
para[51:55,1]=30;   para[51:55,2]=30;   para[51:55,3]=1;    para[51:55,4]=MU2
para[56:60,1]=30;   para[56:60,2]=70;   para[56:60,3]=1;    para[56:60,4]=MU2
para[61:65,1]=30;   para[61:65,2]=10;   para[61:65,3]=4;    para[61:65,4]=MU2
para[66:70,1]=30;   para[66:70,2]=30;   para[66:70,3]=4;    para[66:70,4]=MU2
para[71:75,1]=30;   para[71:75,2]=70;   para[71:75,3]=4;    para[71:75,4]=MU2
para[76:80,1]=30;   para[76:80,2]=10;   para[76:80,3]=9;   para[76:80,4]=MU2
para[81:85,1]=30;   para[81:85,2]=30;   para[81:85,3]=9;   para[81:85,4]=MU2
para[86:90,1]=30;   para[86:90,2]=70;   para[86:90,3]=9;   para[86:90,4]=MU2
para[91:95,1]=70;   para[91:95,2]=10;   para[91:95,3]=1;    para[91:95,4]=MU2  
para[96:100,1]=70;  para[96:100,2]=30;  para[96:100,3]=1;   para[96:100,4]=MU2
para[101:105,1]=70; para[101:105,2]=70; para[101:105,3]=1;  para[101:105,4]=MU2
para[106:110,1]=70; para[106:110,2]=10; para[106:110,3]=4;  para[106:110,4]=MU2
para[111:115,1]=70; para[111:115,2]=30; para[111:115,3]=4;  para[111:115,4]=MU2
para[116:120,1]=70; para[116:120,2]=70; para[116:120,3]=4;  para[116:120,4]=MU2
para[121:125,1]=70; para[121:125,2]=10; para[121:125,3]=9; para[121:125,4]=MU2
para[126:130,1]=70; para[126:130,2]=30; para[126:130,3]=9; para[126:130,4]=MU2
para[131:135,1]=70; para[131:135,2]=70; para[131:135,3]=9; para[131:135,4]=MU2

set.seed(1234) #Seeded for Replication
for(i in 1:135){
  sigma1 = para[i,3];  #Search Para for variance1
  sigma2 = 1; #Search Para for variance2
  n1 = para[i,1]; #Search Para for n1
  n2 = para[i,2]; #Search Para for n2
  mu1 = 5; #Search Para for mean1
  mu2 = para[i,4]; #Search Para for mean2

  #for loop to go through each case  
  #n1 is 1st column of para matrix, n2 is 2nd column of para matrix, sigma1 is 3rd column of para matrix and    mu1 is 4th column of para matrix
  #sigma2 will always equal 1 and mu1 will always equal 5
  
  for(b in 1:B){
    Y1 = rnorm(n1, mu1, sqrt(sigma1)) #Creates randomly generated values from a normal distribution
    Y2 = rnorm(n2, mu2, sqrt(sigma2)) #Creates randomly generated values from a normal distribution
    
    #Pooled
    Sp_2 = ((n1-1)*sigma1^2 + (n2-1)*sigma2^2)/(n1+n2-2); #equal variance
    T_1 = (mu1-mu2)/(sqrt(Sp_2)*sqrt(1/n1 + 1/n2)) #equal variance test statistic
    df_1 = n1 + n2 - 2 #Degrees of Freedom
    pvalue1 = 2*(1-pt(T_1, df_1)) #equal variance p-value
    
    #Satterthwaite
    T_2 = (mu1-mu2)/(sqrt((sigma1^2/n1)+(sigma2^2/n2))) #Test Statistic for Satterthwaite
    df_2 = ((sigma1^2/n1)+(sigma2^2/n2)^2)/((((sigma1^2/n1)^2)/(n1-1) + (((sigma2^2/n2)^2)/(n2-1)))) #Degrees of Freedom for the Satterthwaite t-test unpooled.  
    pvalue2 = 2*(1-pt(T_2, df_2)) #unequal variance p-value
    
    if(pvalue1 <= alpha){  # checking if p-value for equal variance test is less than or equal to alpha 
      result1[i, b] = 1    # recording number of rejections where ith row corresponds to ith situation/combination/scenario
    } else{                # if p-value is greater than alpha
      result1[i, b] = 0    # recording number of fail to rejections 
    }
    if(pvalue2 <= alpha){  # testing if p-value for unequal variance test is less than or equal to alpha
      result2[i, b] = 1    # recording number of rejections
    } else{                # if p-value is greater than alpha
      result2[i, b] = 0    # recording number of fail to rejections
    }
  }
  
}

as.data.frame(Y1)
ggplot(data=Y1, mapping=aes()) +
  geom_histogram()
  

dim(Type1error_1) <- c(27,5)

Type1error_1 <- rowMeans(result1)  #this is type1error for the equal variance test and is the mean/proportion of results recorded of 100 samples 
Type1error_2 <- rowMeans(result2)  #this is type1error for the unequal variance test and is the mean/proportion of results recorded of 100 samples 

tabyl(Type1error_1) #Breaks down counts of Rejects and FAil to Rejects as a proportion.
tabyl(Type1error_2) #Breaks down counts of Rejects and FAil to Rejects as a proportion.
```